### 2016-01-02 pdocs/thesis/main.tex

% Simo.Nikula@gmail.com
% based on latex template at http://www.lut.fi/web/en/library/dissertations1 
\def\mytitle{Handling of plasticity in physics engine}
\def\myisbn{978-952-265-XXX-X}
\def\mypdfisbn{978-952-265-XXX-X}
\def\myyear{2017}
\def\mymonth{May}
\def\myname{Simo Nikula}
\documentclass[a4paper,twoside,12pt,notitlepage,openright]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{lastpage}   % page count
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[english]{babel}
\usepackage[intoc]{nomencl} % including table of content
\usepackage{ifthen}
\usepackage[hmargin=3.0cm,vmargin=3.6cm]{geometry} % setting marginals
\usepackage{fancyhdr,extramarks}  % header ja footer manipulation
\usepackage{times}  % to change font to times
\usepackage{setspace} % for linespacing

\newcommand{\nomunit}[1]{\renewcommand{\nomentryend}{\hspace*{\fill}#1}} % Inserts units on the right at symbol list
\renewcommand{\nomgroup}[1]{%
 \ifthenelse{\equal{#1}{C}}{\item[\textbf{Latin alphabet}]\item}{%
 \ifthenelse{\equal{#1}{G}}{\item\item[\textbf{Greek alphabet}]\item}{}}{%
 \ifthenelse{\equal{#1}{L}}{\item\item[\textbf{Subscripts}]\item}{}}{%
 \ifthenelse{\equal{#1}{H}}{\item\item[\textbf{Superscripts}]\item}{}}{%
 \ifthenelse{\equal{#1}{W}}{\item\item[\textbf{Abbreviations}]\item}{}}{}}

\singlespacing

\pagestyle{fancy}
\fancyhf{}% clearing the header and footer
\fancyhead[LE,RO]{\bfseries\thepage}    % page number to header
\fancyhead[LO]{\nouppercase{\bfseries\rightmark}}     %
\fancyhead[RE]{\nouppercase{\bfseries\leftmark}}%

\renewcommand\sectionmark[1]
 {\markboth{\thesection\ #1}{}}         % section name to header
\renewcommand\subsectionmark[1]
 {\markright{\thesubsection\ #1}}       % subsection name to header
\renewcommand{\headrulewidth}{0.5pt}    % ruler thickness between head and body
\renewcommand{\footrulewidth}{0pt}      % no ruler between body and footer

\setlength{\nomitemsep}{-\parsep}   % removing default extra skip between entries at nomenclature

\numberwithin{equation}{section}    % equation numbers with section numbers
\numberwithin{table}{section}       % table numbers with section numbers
\numberwithin{figure}{section}      % figure numbers with section numbers

% makeindex command needs to run at command prompt to create nomenclature list file
\makenomenclature % makeindex main_v2.nlo -s nomencl.ist -o main_v2.nls
%\bibpunct{(}{)}{;}{a}{,}{,}%

\begin{document}

\pagestyle{empty}
\include{nimio}
\include{abstract}
\addtocontents{toc}{\contentsline {section}{Abstract}{}}

\include{acknowledgments}
\addtocontents{toc}{\contentsline {section}{Acknowledgments}{}}%
\input{preface}
\cleardoublepage%
\thispagestyle{empty}%
\addtocontents{toc}{\contentsline{section}{Contents}{}}%

\tableofcontents%
\thispagestyle{empty}%
\cleardoublepage%
\pagestyle{fancy}%

\cleardoublepage%
\printnomenclature[2.0cm]%

\include{introduction}

%\def\refname{what ever}% to change References title
\addcontentsline{toc}{section}{References}%

\fancyhead[LO]{\nouppercase{\bfseries\firstleftxmark}}%
\fancyhead[RE]{\nouppercase{\bfseries\lastrightxmark}}%
\extramarks{}{References}   % this will remove the "References" to be appearing at the header at first page of bibliography
\extramarks{References}{References}
\bibliography{ref}%
\bibliographystyle{LUTapa}%

%\cleardoublepage %

\end{document}