### 2016-01-03 pdocs/thesis/discussion.tex
pdocs

\section{Discussion}

\subsection{Future work}
\label{subsec:futureWork}

\subsubsection{Large displacements}

\subsubsection{Material nonlinearity}

\cleardoublepage