### 2016-01-07 pdocs/thesis/nimio.tex
pdocs

\thispagestyle{empty} \setlength{\parindent}{0pt}
\begin{figure}
\includegraphics[width=65mm]{./figs/Merkki_Logo_CMYK}\\
\end{figure}
~\\

\vspace{60mm}

{\sffamily\large Simo Nikula\\
\\
\MakeUppercase{\Large \mytitle
}}\\

%\vspace{4mm}
{\sffamily\footnotesize Thesis for the degree of Doctor of Science (Technology) to be presented with\\
due permission for public examination and criticism in the Auditorium ...\\
at \lut, Lappeenranta, Finland on the\\
..th of \mymonth, \myyear, at noon.}\\

\vspace{\stretch{1}}
{\sffamily\normalsize Acta Universitatis\\
Lappeenrantaensis XXX}

\pagebreak

\thispagestyle{empty}

\begin{tabular}{l p{11.0cm}}Supervisor& Professor Timo Björk\\
&Laboratory of Steel Structures\\
&School of Energy Systems\\
&Lappeenranta University of Technology\\
&Finland\\
\\
Reviewers& Title Name\\
&Department of \\
&University of \\
&Country\\
\\
& Title Name\\
&Department of \\
&University\\
&Country\\

\\
Opponent& PhD Name\\
&Institution\\
&Country\\
\end {tabular}

\vspace{\stretch{1}}

\begin{center}
ISBN \myisbn\\
ISBN \mypdfisbn (PDF)\\
ISSN 1456-4491\\
Lappeenrannan teknillinen yliopisto\\
Yliopistopaino \myyear

 \end{center}
\pagebreak