### 2016-01-09 pdocs/thesis/discussion.tex
pdocs

\section{Discussion}

\subsection{Future work}
\label{subsec:futureWork}

\subsubsection{Acceptance to current physics engines}

Required work depends on selected dynamics engine. Availability of these features makes similar 
implementation possible.

\begin{itemize}
\item generic spring or motor constraint is already available 
\item possibility to update constraints after each simulation step
\end{itemize}

Current implementation is done based on \bullet but it still needs refactoring and testing before it can be 
accepted to main stream. Interfaces should be reafctored so that handling of additional features like large 
displacements and material nonlinearity can be implemented without further changes.

\subsubsection{Large displacements}
Large displacements are typical if plasticity is involved and
large displacements can also take place due to rigid body motion and should be taken into account.

Update of constraints can be done in  many ways but using callbacks allows clean integration i.e. no changes
to solver core are needed.
In \bullet there are already three per step callbacks.

\begin{itemize}
\item updateActions
\item updateActivationState
\item btInternalTickCallback
\end{itemize}

\demolisher already uses updateActions and that seems 
good way to do also other updates. Special care should be taken to avoid unnecessary updates.

\subsubsection{Material nonlinearity}
Update of constraint properties can be done in callbacks in similar way as for large displacements.
Material definition should be defined so that stress-strain curve can be defined as array of pairs.

\cleardoublepage