### 2016-09-19 pdocs/thesis/article-conclusion.tex
skeleton for conclusion

\section{Conclusion}

This study has presented an approach to account plastic deformation in 
velocity based formulation.
In the introduced method, the plastic deformation takes place if force or moment exceeds given 
limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
Approach presented in this work can be used in gaming industry to provide more realistic 
simulations without significant extra work. For gaming purposes presented method works 
best in scenarios where connected parts are relatively heavy. This allows normal 
integration timestep to be used without stability issues. 
This kind of metodology also opens large area of combining old structural analysis
methods to modern simulation frameworks.
As already noted in \cite{cg1988}, the modeling of inelastic deformation
remains open for further exploration in the context of computer graphics.