### 2016-10-08 pdocs/thesis/abstract.tex
Aki's comments on article and in general

\section*{Abstract}

\textbf{\myname}\\
\textbf{\mytitle}\\
Lappeenranta \myyear\\
\pageref{LastPage} pages\\
Acta Universitatis Lappeenrantaensis XXXXXX\\
Diss. Lappeenranta University of Technology\\
ISBN \myisbn, ISBN \mypdfisbn (PDF), ISSN 1456-4491\\

In this work suitable methods for handling of plasticity in physics engine are presented.
Handling of plasticity basically means that involved bodies may absorb energy and possibly break.

Methods are verified by extending Bullet Physics Library.

Keywords: Physics engine, simulation, dynamics, plasticity, inelastic, multibody dynamics,  physical simulation, deformation.