### 2016-10-29 pdocs/thesis/article-section-2.tex
working on Peter's comments

% !TeX root = article.tex
\section{Description of plasticity in the framework of physics engines}

In this section, key concepts related to the introduced model are explained. The main differences between 
traditional structural analysis and physics engines based approaches are reviewed and discussed.

Velocity-based formulation of constraint based rigid body simulation 
is  commonly used by physics based game
developers and film production teams.
%\citet[p.~45]{erleben.thesis} 
\citet{erleben.thesis} 
provides reasoning and theoretical details for the popularity of 
velocity-based formulation in constraint-based rigid body simulation instead of accelaration based. 
The main reason is that collision handling can be done without the use of additional procedures.

Work presented by  
\citet{erleben.thesis} provides the basis for the velocity-based formulation discussed in this work.
%\citet[p.~45-50]{erleben.thesis}.
% pdf page 64
In the following section, these formulations will be clarified by a simple example using \bullet\ implementation.

Impulse $\vec{J}$
in the time interval $\Delta t $ can be written as
\begin{equation} \label{eq:impulseIntegral}
\vec{J} = \int_{0}^{\Delta t} \vec{f}_{true}(t) dt,
\end{equation}
where $\vec{f}_{true}(t)$ is force.

Using Newton's second law of motion $\vec{F}=m\vec{a}$ ,
$\vec{v}^{\Delta t}$ can be solved as
\begin{equation} \label{eq:impulseIntegraWithNewton}
\int_{0}^{\Delta t} m \frac{d\vec{v}}{dt}dt= \int_{0}^{\Delta t} \vec{f}_{true}(t)
\end{equation}
\begin{equation} \label{eq:impulse}
m(\vec{v}^{\, \Delta t} - \vec{v}^{\, 0})=\vec{J},
\end{equation}
where superscripts denote time, i.e. ${\vec{v}}^{\Delta t}=\vec{v}(\Delta t)$.
Next position can be found
by integrating the velocity.
Updates after each step can be summarized  for locations and  
for velocities respectively as follows:

\begin{equation} \label{eq:eomL} % pdf page 69
\vec{s}^{\, t+\Delta t} = \vec{s}^{\, t}+\Delta t S \vec{u}^{\, t+\Delta t}
\end{equation}

\begin{equation} \label{eq:eomV}
\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(C N \vec{f}^{\ t+\Delta t} + \vec{f}_{ext}) .
\end{equation}

The symbols used in Equations \ref{eq:eomL} and \ref{eq:eomV}
are summarized in Table \ref{tab:eom}.
Figure \ref{fig:eom-contact} describes the collision of two bodies, $B_1$ and $B_2$
where $\vec{r}_i$ is orientation of body $i$,
$\vec{p}_k $ is orientation of contact point $k$,
$\vec{r}_{ik} $ is a vector  between the center of gravity of body $i$ and contact point $k$, 
and
$\vec{n}_{k}$ is the contact normal for contact point $k$.

\begin{figure}[tb!]
\centering
\begin{tikzpicture}
\coordinate (O1) at(2,1);
\coordinate (O2) at(2,4);
\coordinate (C) at(2,2);
\draw (0,0) -- (4,0) -- (4,2) -- (0,2) --(0,0);
\draw (2,2) -- (3.5,5) -- (0.5,5) -- (2,2) ;
\node at (3.5,0.4) {$B_1$};
\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
\node at (2.8,4.5) {$B_2$};
\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
\filldraw (C) circle (0.5mm) node[anchor=north west] {$\vec{p}_1$};
\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{11}$};
\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{21}$};
\draw[-latex,thick] (C) -- ++(0,1.4) node[anchor=west] {$\vec{n}_{1}$};
\node[anchor=west] at (4.5,4.5) {
$\vec{r}_i$  =  orientation of body $i$
};
\node[anchor=west] at (4.5,4) {$\vec{p}_k $ = orientation of contact point $k$};
\node[anchor=west] at (4.5,3.5) {$\vec{r}_{ik} $ = $\vec{p}_k - \vec{r}_i $};
\node[anchor=west] at (4.5,3) {$\vec{n}_{k} $ = normal for contact point $k$};
\end{tikzpicture}
\caption{Illustration of nomenclature for equations of motion for collision.}
\label{fig:eom-contact}
\end{figure}

% pdf page 33, notation in typical ODEs
\begin {table}[htb!]
\caption {Nomenclature for equations of motion}
\label{tab:eom}
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$\vec{r}_i$ & position of center of mass for body $i$  \\ \hline
$\vec{q}_i$ & orientation for body $i$ as quaternion $\lbrack s_i, x_i, y_i, z_i \rbrack ^T $ \\
\hline
$\vec{p}_h$ & contact or joint point $k$  \\ \hline
$\vec{r}_{ki}$ & $\vec{p}_k - \vec{r}_i$  \\ \hline
$\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
$Q_i$ & \begin{tabular}{@{}c}
rotation of quaternion $\vec{q}_i$
as matrix \\ where 
$\frac{1}{2}\vec{\omega}_i \vec{q}_i=Q_i \vec{\omega}_i$
\end{tabular}
   $
\frac{1}{2} \left[ \begin{array}{ccc}
-x_i & -y_i & -z_i \\
s_i & z_i & -y_i \\
-z_i & s_i & x_i \\
y_i & -x_i & s_i
\end{array} \right]
$
\\ \hline
$S$ & 
\begin{tabular}{@{}c}
generalized transformation matrix \\
$ S \in \mathbb{R}^{7n \times 6n}$
\end{tabular}
 $ \left[ \begin{array}{ccccc}
1 &  &  & & 0 \\
 & Q_i  \\
 & & \ddots  \\
 & & & 1 \\
0 & & & & Q_n 
\end{array} \right]
$
\\ \hline
$\vec{v}_i$ & linear velocity of  center of mass for body $i$   \\ \hline
$\vec{\omega}_i$ & angular velocity of center of mass for body $i$  \\ \hline
$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
$M$ &
\begin{tabular}{@{}c}
 generalized mass matrix \\
$ M \in \mathbb{R}^{6n \times 6n}$  
\end{tabular}
$
\left[ \begin{array}{ccccc}
m_i 1 &  &  & & 0 \\
 & I_1  \\
 & & \ddots  \\
 & & & m_n 1 \\
0 & & & & I_n 
\end{array} \right]
$
\\ \hline
$I_i$ & inertia tensor for body $i$  \\ \hline
$C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$ \\ \hline
\end {tabular}
\end{center}
\end {table}

Friction in contacts and joint constraints can be handled in a unified way by refactoring
equation \ref{eq:eomV} as,  
\citet{erleben.thesis}
%\citet[p.~66-67]{erleben.thesis}
\begin{equation} \label{eq:eomV2}
\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(
 J_{contact}^T \vec{\lambda}_{contact}
+ J_{joint}^T \vec{\lambda}_{joint}
+ \vec{f}_{ext}),
\end{equation}
where Jacobian terms $J_{contact}^T$ for joints are 
derived by taking time derivatives of the kinematic constraints.

Symbols used in Equation \ref{eq:eomV2} are summarized in Table
\ref{tab:eom-g} and Figure \ref{fig:eom-joint}.
Figure \ref{fig:eom-joint} presents terms needed for joint processing.

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (O1) at(1,1);
\coordinate (O2) at(1,3);
\coordinate (C) at(1,2);
\draw (0,0) -- (2,0) -- (2,2) -- (0,2) --(0,0);
\draw (0,2) -- (2,2) -- (2,4) -- (0,4) --(0,2) ;
\node at (0.5,0.4) {$B_1$};
\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
\node at (0.5,3.5) {$B_2$};
\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{anc}^{\,1}$};
\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{anc}^{\,2}$};
\node[anchor=west] at (4.5,3.5) {
$\vec{r}_i$  =  orientation of body $i$ 
};
\node[anchor=west] at (4.5,3) {$\vec{r}_{anc}^{\,i} $ = body frame vector $i$};
\end{tikzpicture}
\caption{Illustration of nomenclature for equations of motion for joint.}
\label{fig:eom-joint}
\end{figure}

In Figure \ref{fig:eom-joint}, $\vec{r}_{anc}^{\,i}$ is used to define at which point
joint constraint is applied.

% pdf page 33, notation in typical ODEs
\begin {table}[htb!]
\caption {Additional terms  for generalized equations of motion}
\label{tab:eom-g}
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
$J_{joint}$ & Jacobian matrix for joints  \\ \hline
$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
\end {tabular}
\end{center}
\end {table}

Constraint processing in \bullet\ is based on ODE, \cite{ode}.
Joints are also discussed in detail in  
\citet{erleben.thesis}.
%\citet[p.~60-90]{erleben.thesis}.
Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and
\ref{eq:lambdaHigh} 
are created for each constraint. 
Derivation for terms in Equation \ref{eq:constraintEquation}
can be done using the position and orientation of connected bodies
e.g. for ball joint formulation is based on both joint points having the same position.
In contact cases, formulation is easier if it is done using velocities, \cite{ode.joints}.

\begin{equation} \label{eq:constraintEquation}
J_1 \vec{v}_1 + \Omega_1 \vec{\omega}_1 + 
J_2 \vec{v}_2 + \Omega_2 \vec{\omega}_2 = \vec{c} + C \vec{\lambda}
\end{equation}

\begin{equation} \label{eq:lambdaLow}
\vec{\lambda} \geq \vec{l}
\end{equation}

\begin{equation} \label{eq:lambdaHigh}
\vec{\lambda} \leq \vec{h}
\end{equation}

In the following section, these equations will be explained by a simple example.
The main parameters  and corresponding fields in \bullet\  
are given in Table \ref{tab:constraintParameters}.

\begin {table}[htb!]
\caption {Constraint parameters}
\label{tab:constraintParameters} 
\begin{center}
\begin{tabular}{|c| l| l|}
\hline
{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\
$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
$\vec{v}$ & linear velocity & \\ \hline
$\vec{\omega}$ & angular velocity & \\ \hline
$\vec{c}$        &  right side vector   & m\_constraintError \\ \hline
$C$  & constraint force mixing & cfm \\  \hline
$\vec{\lambda}$ & constraint force &  \\ \hline
$\vec{l}$ & low limit for constraint force & m\_lowerLimit \\ \hline
$\vec{h}$ & high limit for constraint force & m\_upperLimit \\ \hline
\end {tabular}
\end{center}
\end {table}

In structural analysis, a formulation and associated numerical solution procedure are selected 
based on needed features.
Often,  finite element method is used.
In most cases, a static solution with an assumption of linear strain-displacement relation
using displacement based boundary conditions is used.
\citet{bathe-1975} provides a description for handling of various nonlinearities.
In large displacement analysis, formulation may be based on updated formulation (Eulerian) or
Lagrangian formulation where initial configuration is used.
Further enhancements are material nonlinearity and dynamic analysis.
Physics engine provides dynamic analysis with large reference translations and rotations
while assuming bodies to be undeformable.

Material plasticity can be accounted for in simulations by using a suitable coefficient of restitution.
This provides a reasonable means to simulate loss of energy in collisions.
In this work simulation of breaking of bodies made of ductile material is made more realistic 
by splitting the rigid body
to multiple bodies that are connected by energy absorbing joints.
A typical engineering stress-strain curve of ductile steel is shown in Figure \ref{fig:sscurve}.

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (Y) at (1,4);
\draw[->] (0,0) -- (10,0) node[right] {\large{$\epsilon$}};
\draw[->] (0,0) -- (0,6) node[above] {\large{$\sigma$}};
\draw(0,0) -- (Y) -- (2,4) .. controls (7,6) .. (10,5);
\draw[dashed](0,4) -- (Y);
\node at (-0.2,4) [align=right] {$f_y$};
\draw(0.25,1) -- (0.5,1) -- (0.5,2);
\node at (0.75,1.5) {$E$};
\node at (0.8,2.5) [anchor=west] {$\sigma = E \epsilon$ if $\sigma \le f_y$};
\end{tikzpicture}
\caption{Engineering stress-strain curve of ductile steel (not to scale).}
\label{fig:sscurve}
\end{figure}

In Figure \ref{fig:sscurve}, $\sigma$ is stress, $E$ is Youngs modulus and $f_y$ is yield stress.
Engineering stress and strain mean that original dimensions are used in stress calculation,
\citet{dowling}.
%\citet[p.~108]{dowling}.
The stress-strain curve is not drawn to scale as elastic strain could not be seen as it is typically 
0.001 to 0.005 and fracture strain can be 100 times larger.

In this work, an elastic-fully plastic material model is used in most scenarios.
Having elastic part allows elastic displacements for slender structures. 
Elastic material behavior is ignored in approach introduced in this work if
the deformation is related to ahigher frequency
than integration stability would allow.
It should be noted that geometry
of bodies is not updated during analysis and thus engineering stress-strain properties are used.

In this work, strain hardening is taken into account by assuming that plastic volume in bending
expands, 
\citet{dowling}.
%\citet[p.~672]{dowling}.
Material that starts to yield first is hardened and as a result of which yielding moves.

The difference between the elastic and plastic section modulus is depicted in Figure \ref{fig:wp}.

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (S) at (2.5,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1.5,0) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma<f_y$}};
\node[anchor=west] at (-2,3) {
\begin{tabular}{l}
Under elastic load\\
stress increases\\
linearly from zero\\
at neutral axis to\\
maximum value at \\
surface of body
\end{tabular}
};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}
\coordinate (S) at (3,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1,0) -- (1,2.5) -- (3,2.5) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma=f_y$}};
\node[anchor=west] at (3,3) {
\begin{tabular}{l}
Under fully plastic load\\
stress is at yield\\
level over full\\
cross section
\end{tabular}
};
\end{tikzpicture}
\caption{Axial stress distribution over a cross section for bending under elastic and fully plastic loads.}
\label{fig:wp}
\end{figure}

As shown in Figure 2.4, if stress is below yield limit $f_y$, stress and strain are linear within the material.
If cross section is fully plastic, stress is assumed to be at yield level over the whole cross section such that 
the plastic section modulus is higher than the elastic section modulus.

In this work, plasticity is handled by defining maximum forces
in Equations \ref{eq:lambdaLow} and  
\ref{eq:lambdaHigh} using plastic capacities, which are defined below.

Maximum force acting in a direction of $\vec{r}_{anc}^{\,i} $
is product of area and yield stress as follows

\begin{equation} \label{eq:fN}
N_{max}= \int_A f_y
\end{equation}

Maximum forces acting perpendicular to $\vec{r}_{anc}^{\,i} $
are a product of area and shear yield stress $\tau_y$ as follows
\begin{equation} \label{eq:fQ}
Q_{max}= \int_A \tau_y
\end{equation}

Maximum moments acting around the axis perpendicular to $\vec{r}_{anc}^{\,i} $
are integrals of the perpdendicular distance 
and yield stress $f_y$ as given for the moment around the $x$-axis 
and moment around the $z$-axis, respectively.

\begin{equation} \label{eq:Mx}
M_{max}^x= \int_A z f_y
\end{equation}

\begin{equation} \label{eq:Mz}
M_{max}^z= \int_A x f_y
\end{equation}

Maximum moment around $\vec{r}_{anc}^{\,i} $
is an integral of distance $d$ from the joint point
and shear yield stress $\tau_y$ as

\begin{equation} \label{eq:My}
M_{max}^y= \int_A d \tau_y
\end{equation}

Maximum forces and moments for a
rectangular section with width $b$ and height $h$ using constant yield stress
are given in Table \ref{tab:maxForces}.
Yield shear stress is assumed to be $ 0.5\, f_y$ using the Tresca yield critetion.
If the von Mises yield criterion is used 0.5 is replaced by 0.58 ($1/\sqrt{3}$), \cite{dowling}.
% p. 262, p. 268
These are not exact values in a multiaxial stress state but they
should be acceptable in most gaming scenarios.

\begin {table}[htb!]
\caption {Maximum forces and moments for 
rectangular section with width $b$ and height $h$ using constant yield stress $f_y$}
\label{tab:maxForces} 
\begin{center}
\begin{tabular}{| c| c|}
\hline
{\bf Direction} & {\bf Maximum value}  \\ \hline
maximum shear force & $0.5\, b\, h f_y$ \\ \hline
maximum normal force & $b\, h\, f_y$  \\ \hline
maximum bending moment in direction of $h$& $0.25\, b\, h^2 \, f_y$  \\ \hline
maximum bending moment in direction of $b$ & $0.25\, b^2\, h\, f_y$  \\ \hline
maximum torque & $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\ \hline
\end{tabular}
\end{center}
\end {table}

For torque there is a closed form solution only for
circular cross sections.
Given approximation is 
best suited for cases where $b$ and $h$ are similar.
Better approximation for any given $b$ and $h$ can be obtained 
by integrating distance from the center of the joint over cross section and
multiplying it with the yield shear stress e.g. using Octave, \cite{octave}.
An example of calculation of the maximum moment  around $\vec{r}_{anc}^{\,i} $
is shown in Figure \ref{fig:octave-mp}

\begin{figure}[htb!]
\centering
\lstset{language=octave}
\begin{lstlisting}
b=0.01; h=0.01; fy=200e6;
wpy=fy/2*dblquad(@(x,z) sqrt(x.*x+z.*z),-b/2,b/2,-h/2,h/2)
38.2
\end{lstlisting}

\caption{Calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $ using Octave.}
\label{fig:octave-mp}
\end{figure}

The basic idea introduced in this study can be tested with any framework having motors and hinge constraints.
This can be done by setting the target velocity of the motor to zero and limiting 
the maximum motor impulse to plastic moment multiplied by a timestep.

Further enhancements were created and tested by forking \bullet\ source code
and adding new constraints, \cite{pbullet}.
Instructions for using  windows executable and  source code are available, \cite{bp}.