### 2016-10-31 pdocs/thesis/article.tex
acm tog styling of article continues

% Simo.Nikula@gmail.com
\def\mytitle{Elastic-plastic Constraint for Rigid Body Simulation}
\def\lut{Lappeenranta University of Technology}
\def\bullet{Bullet Physics}
\def\demolisher{DemolisherDemo}
\def\charpy{CharpyDemo}

\documentclass{acmtog}
\acmVolume{VV}
\acmNumber{N}
\acmYear{201Y}
\acmMonth{Month}
\acmArticleNum{AN}  
\acmdoi{10.1145/xxx.yy}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
%\usepackage{natbib}
%\usepackage{lastpage}   % page count
\usepackage{amsmath}
\usepackage{listings}
% \usepackage{mathdots}  % improved \ddots and \iidots
\usepackage{amssymb}
%\usepackage[english]{babel}
%\usepackage[intoc]{nomencl} % including table of content
%\usepackage{ifthen}
%\usepackage[hmargin=2.0cm,vmargin=1.6cm]{geometry} % setting marginals
%\usepackage{fancyhdr,extramarks}  % header ja footer manipulation
%\usepackage{times}  % to change font to times
%\usepackage{setspace} % for linespacingx
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta}
%\usepackage{caption}
\usepackage{varwidth}
\usepackage{enumitem} % 
\begin{document}
\markboth{S. Nikula et al.}{\mytitle}

\title{\mytitle}

\author{SIMO NIKULA {\upshape and} AKI MIKKOLA {\upshape and} TIMO BJÖRK
\affil{\lut}}

\category{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}[Animation]
\category{I.6.3}{Computer Graphics}{Simulation and Modelling}[Applications]

\keywords{Modeling, Animation, Deformation, Elasticity, Plasticity, Dynamics, Simulation}

\acmformat{Nikula, S., Mikkola, A., Björk, T.,  G. V. G. 201x. 
\mytitle.  {ACM Trans. Graph.} VV, NN, Article AN (Mon 201x), nb pages.\newline  DOI $=$
10.1145/xx.yy\newline http://doi.acm.org/10.1145/XXXXXXX.YYYYYYY}

\maketitle

\begin{bottomstuff}

Authors' addresses: Simo.Nikula@gmail.com
\end{bottomstuff}

\input{article-introduction}
\input{article-section-2}
\input{article-section-3}
\input{article-conclusion}

\bibliographystyle{acmtog}
\bibliography{ref}

\end{document}