### 2016-11-01 pdocs/thesis/article.tex

% Simo.Nikula@gmail.com
\def\mytitle{Elastic-plastic Constraint for Rigid Body Simulation}
\def\lut{Lappeenranta University of Technology}
\def\bullet{Bullet Physics}
\def\demolisher{DemolisherDemo}
\def\charpy{CharpyDemo}

\documentclass{acmtog}
\acmVolume{VV}
\acmNumber{N}
\acmYear{201Y}
\acmMonth{Month}
\acmArticleNum{AN}  
\acmdoi{10.1145/xxx.yy}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta}
\usepackage{varwidth}
\usepackage{enumitem} % 
\begin{document}
\markboth{S. Nikula, A. Mikkola and T. Björk}{\mytitle}

\title{\mytitle}

\author{SIMO NIKULA {\upshape and} AKI MIKKOLA {\upshape and} TIMO BJÖRK
\affil{\lut}}

\category{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}[Animation]
\category{I.6.3}{Computer Graphics}{Simulation and Modelling}[Applications]

\keywords{Modeling, Animation, Deformation, Elasticity, Plasticity, Dynamics, Simulation}

\acmformat{Nikula, S., Mikkola, A., Björk, T.,  G. V. G. 201x. 
\mytitle.  {ACM Trans. Graph.} VV, NN, Article AN (Mon 201x), nb pages.\newline  DOI $=$
10.1145/xx.yy\newline http://doi.acm.org/10.1145/XXXXXXX.YYYYYYY}

\maketitle

\begin{bottomstuff}

%Authors' addresses: Simo.Nikula@gmail.com
\end{bottomstuff}

\input{article-introduction}
\input{article-section-2}
\input{article-section-3}
\input{article-conclusion}

\bibliographystyle{acmtog}
\bibliography{ref}

\end{document}