### 2017-06-10 pdocs/thesis/epj-jcgt.tex
jcgt article started

% Simo.Nikula@gmail.com
\def\mytitle{Breakable Elastic-plastic Constraint for Rigid Body Simulation}
\def\lut{Lappeenranta University of Technology}
\def\

\documentclass{jcgt}

\setciteauthor{Simo Nikula, Aki Mikkola and Timo Björk}
\setcitetitle{\\mytitle}

% Mark submissions with the date of submission using the following line:
\submitted{\today}

% Once an article is accepted accepted, switch to the following line and comment the preceding one. The editor will supply the argument values.
%\accepted{2014-02-07}{2014-02-07}{2014-02-07}{Editor Name}{3}{1}{1}{1}{2014}
%\seturl{http://jcgt.org/published/0003/01/01/}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\newcommand{\cbullet}[0]{Bullet Physics}
\title{\\mytitle}

\author
       {SIMO NIKULA {\upshape and} AKI MIKKOLA {\upshape and} TIMO BJÖRK\\\lut
       }

\teaser{
  \includegraphics[width=2in]{figs/demolisher-done.PNG}
  \label{fig:teaser}
  \caption{Adding ductile breakable constraint to popular open source physics engines \cbullet\ and Box2D
allows more realistic simulation of many scenarios.
}
}

\maketitle
\thispagestyle{firstpagestyle}

\begin{abstract}
\small
This paper introduces simple and efficient method to simulate ductile fracture in existing 
impulse-based physics engines.
Method is based on technique of splitting bodies to multiple pieces and joining them with constraints.
Constraints are described using body dimensions and material parameters.

Sample programs with source code are made available to allow developers already using \cbullet\
and Box2D to add plasticity into their simulations.
\end{abstract}

%-------------------------------------------------------------------------
\section{Introduction}
\label{sec:introduction}

Theory for handling of plasticity in computer graphics has been presented already 1988, \cite{cg1988}. 
Recent paper by Jones  provides extensive listing of related work during past decades, \cite{Jones:2016:EPD}.

In impulse-based phycis engines most breakable scenarios  are implemented by destructing various bodies based on collision
or impulse exceeding predefined limit.
Nevertheless, breaking of steel or reinforced concrete structures using this approach 
is not appropriate if the simulation is to look realistic.

More realistic simulation of ductile destructable bodies is possible but in many cases it would require selecting of new physics engine.

\paragraph{Our approach}
This study will introduce an approach to account for plastic deformation in impulse-based phycis engines.   
Presented methodology does not require significant software development efforts from
game vendors and is thus easily adoptable. 
In the introduced method, plastic deformation takes place if the force or moment exceeds a predefined 
limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
Maximum forces and moments are estimated based on the plastic section modulus.
Joint breaking is based on summing plastic deformation and comparing it to a
predefined material based limit. The elastic part of deformation is modelled by employing 
modification of an existing constraints.

\paragraph{Limitations}
Forces are calculated by dividing impulse by timestep and maximum impulse is calculated
by multiplying maximum force by timestep. 
This means that simulation of quick interactions requires small timestep.
This limitation is not introduced by this approach but is build in feature of impulse-based simulation.

Multiaxial stress is not taken into account.

\section{Description of plasticity in the framework of physics engines}

In this work simulation of breaking of bodies made of ductile material is made more realistic 
by splitting the rigid body
to multiple bodies that are connected by energy absorbing joints.

\paragraph{Stress-strain behaviour of ductile steel}
A typical engineering stress-strain curve of ductile steel is shown in Figure \ref{fig:sscurve}.

\begin{figure}
\centering
\begin{tikzpicture}
\coordinate (Y) at (1,4);
\draw[->] (0,0) -- (6,0) node[right] {\large{$\epsilon$}};
\draw[->] (0,0) -- (0,5) node[above] {\large{$\sigma$}};
\draw(0,0) -- (Y) -- (2,4) .. controls (5,5) .. (6,4);
\draw[dashed](0,4) -- (Y);
\node at (-0.2,4) [align=right] {$f_y$};
\draw(0.25,1) -- (0.5,1) -- (0.5,2);
\node at (0.75,1.5) {$E$};
\node at (0.8,2.5) [anchor=west] {$\sigma = E \epsilon$ if $\sigma \le f_y$};
\end{tikzpicture}
\caption{Engineering stress-strain curve of ductile steel (not to scale).}
\label{fig:sscurve}
\end{figure}

In Figure \ref{fig:sscurve}, $\sigma$ is stress, $E$ is Youngs modulus and $f_y$ is yield stress.
Engineering stress and strain mean that original dimensions are used in stress calculation,
\cite{dowling}.
%\citet[p.~108]{dowling}.
The stress-strain curve is not drawn to scale as elastic strain could not be seen as it is typically 
0.001 to 0.005 and fracture strain can be 100 times larger.
In practice this means that elastic displacement due to axial forces is usually not visible.
Visible elastic displacement is usually due to bending.

In this work, an elastic-fully plastic material model is used in most scenarios.
Having elastic part allows elastic displacements for slender structures. 
Elastic material behavior is ignored in approach introduced in this work if
the deformation is related to a higher frequency
than integration stability would allow.
It should be noted that geometry
of bodies is not updated during analysis and thus engineering stress-strain properties are used.

In this work, strain hardening is taken into account by assuming that plastic volume in bending
expands, 
\cite{dowling}.
%\citet[p.~672]{dowling}.
Material that starts to yield first is hardened and as a result of which yielding moves.

The difference between the elastic and plastic section modulus is depicted in Figure \ref{fig:wp}.

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (S) at (2.5,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1.5,0) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma<f_y$}};
\node[anchor=west] at (3,3) {
\begin{tabular}{l}
Under elastic load\\
stress increases\\
linearly from zero\\
at neutral axis to\\
maximum value at \\
surface of body
\end{tabular}
};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}
\coordinate (S) at (3,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1,0) -- (1,2.5) -- (3,2.5) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma=f_y$}};
\node[anchor=west] at (3,3) {
\begin{tabular}{l}
Under fully plastic load\\
stress is at yield\\
level over full\\
cross section
\end{tabular}
};
\end{tikzpicture}
\caption{Axial stress distribution over a cross section for bending under elastic and fully plastic loads.}
\label{fig:wp}
\end{figure}

As shown in Figure \ref{fig:wp}, if stress is below yield limit $f_y$, stress and strain are linear within the material.
If cross section is fully plastic, stress is assumed to be at yield level over the whole cross section such that 
the plastic section modulus is higher than the elastic section modulus.

\paragraph{Plastic capacities}
In this work, plasticity is handled by defining maximum forces using plastic capacities, which are defined below.

Maximum force acting in a direction of $\vec{r}_{anc}^{\,i} $
is product of area and yield stress as follows:

\begin{equation} \label{eq:fN}
N_{max}= \int_A f_y.
\end{equation}

Maximum forces acting perpendicular to $\vec{r}_{anc}^{\,i} $
are a product of area and shear yield stress $\tau_y$ as follows:
\begin{equation} \label{eq:fQ}
Q_{max}= \int_A \tau_y.
\end{equation}

Maximum moments acting around the axis perpendicular to $\vec{r}_{anc}^{\,i} $
are integrals of the perpdendicular distance 
and yield stress $f_y$ as given for the moment around the $x$-axis 
and moment around the $z$-axis, respectively:

\begin{equation} \label{eq:Mx}
M_{max}^x= \int_A z f_y,
\end{equation}

\begin{equation} \label{eq:Mz}
M_{max}^z= \int_A x f_y.
\end{equation}

Maximum moment around $\vec{r}_{anc}^{\,i} $
is an integral of distance $d$ from the joint point
and shear yield stress $\tau_y$ as:

\begin{equation} \label{eq:My}
M_{max}^y= \int_A d \tau_y.
\end{equation}

Maximum forces and moments for a
rectangular section with width $b$ and height $h$ using constant yield stress
are given in Table \ref{tab:maxForces}.
Yield shear stress is assumed to be $ 0.5\, f_y$ using the Tresca yield critetion.
If the von Mises yield criterion is used 0.5 is replaced by 0.58 ($1/\sqrt{3}$), \cite{dowling}.
% p. 262, p. 268
These are not exact values in a multiaxial stress state but they
should be acceptable in most gaming scenarios.

\begin {table}
\tbl {Maximum forces and moments for 
rectangular section with width $b$ and height $h$ using constant yield stress $f_y$}{
\begin{tabular}{| c| c|}
\hline
{\bf Direction} & {\bf Maximum value}  \\ \hline
maximum shear force & $0.5\, b\, h f_y$ \\ \hline
maximum normal force & $b\, h\, f_y$  \\ \hline
maximum bending moment in direction of $h$& $0.25\, b\, h^2 \, f_y$  \\ \hline
maximum bending moment in direction of $b$ & $0.25\, b^2\, h\, f_y$  \\ \hline
maximum torque & $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\ \hline
\end{tabular}}
\label{tab:maxForces} 
\end {table}

For torque there is a closed form solution only for
circular cross sections.
Given approximation is 
best suited for cases where $b$ and $h$ are similar.
Better approximation for any given $b$ and $h$ can be obtained 
by integrating distance from the center of the joint over cross section and
multiplying it with the yield shear stress e.g. using Octave, \cite{octave}.
An example of calculation of the maximum moment  around $\vec{r}_{anc}^{\,i} $
is shown in Figure \ref{fig:octave-mp}.

\lstset{language=octave}
\begin{lstlisting}[caption={Calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $ using Octave.}, 
label={lst:octave-mp}, float]
b=0.01; h=0.01; fy=200e6;
wpy=fy/2*dblquad(@(x,z)...
 sqrt(x.*x+z.*z),-b/2,b/2,-h/2,h/2)
38.2
\end{lstlisting}

\paragraph{Joint breaking}

\section{Implementation for Box2D}

\section{Implementation for \cbullet\}

\small
\bibliographystyle{jcgt}
\bibliography{paper}

\section*{Index of Supplemental Materials}

\section*{Author Contact Information}

\hspace{-2mm}\begin{tabular}{p{0.5\textwidth}p{0.5\textwidth}}
Simo Nikula\newline
\href{mailto:Simo.Nikula@gmail.com}{Simo.Nikula@gmail.com}
&
Aki Mikkola\newline
\href{mailto:Aki.Mikkola@lut.fi}{Aki.Mikkola@lut.fi}
&
Timo Björk\newline
\href{mailto:Timo.Bjork@lut.fi}{Timo.Bjork@lut.fi}

\end{tabular}

\afterdoc

\end{document}